---
title: "D-Final Data Prep w Met Office and NDVI"
author: "Abi Riley"
date: "2024-12-28"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "Z:/home")

```


## Packages 

```{r packages, echo= FALSE}

library(spatstat)
library(sp)
library(gstat)
library(sf)
library(dplyr)
library(MatrixModels)
library(gtools)
library(rquery)

```

Load data and change coords

```{r data}

load(file = "DATA/Data/C6-All and Sitedata w Road Dens.RData")

head(all)

all[,c("Easting", "Northing")] <- all[,c("Easting", "Northing")]/1000


#mean(all$PM25PCM, na.rm = TRUE)


```


Calculate wind variables

```{r wind}

all$Wind_speed <- sqrt(all$Wind_u^2 + all$Wind_v^2)
wind_trig <- atan2(all$Wind_u/all$Wind_speed, all$Wind_v/all$Wind_speed)
all$Wind_deg <- wind_trig * 180/pi + 180

library(dplyr)
North1 <- c(337.5, 360)
North2 <- c(0, 22.5)
NorthEast <- c(22.5, 67.5)
East <- c(67.5, 112.5)
SouthEast <- c(112.5, 157.5)
South <- c(157.5, 202.5)
SouthWest <- c(202.5, 247.5)
West <- c(247.5, 292.5)
NorthWest <- c(292.5, 337.5)

all$Wind_dir <- ifelse(between(all$Wind_deg, North1[1], North1[2]), "N",
                       ifelse(between(all$Wind_deg, North2[1], North2[2]), "N",
                              ifelse(between(all$Wind_deg, NorthEast[1], NorthEast[2]), "NE",
                                     ifelse(between(all$Wind_deg, East[1], East[2]), "E",
                                            ifelse(between(all$Wind_deg, SouthEast[1], SouthEast[2]), "SE",
                                                   ifelse(between(all$Wind_deg, South[1], South[2]), "S",
                                                          ifelse(between(all$Wind_deg, SouthWest[1], SouthWest[2]), "SW",
                                                                 ifelse(between(all$Wind_deg, West[1], West[2]), "W",
                                                                        ifelse(between(all$Wind_deg, NorthWest[1], NorthWest[2]), "NW", NA)))))))))

count(all$Wind_dir)
head(all)

```

Crop 

```{r crop}

Wards <- read_sf("Jobs/ST_INLA/Shapefiles/London_Ward.shp")
London <- st_transform(Wards, crs = "+proj=tmerc +lat_0=49 +lon_0=-2 +k=0.999601272 +x_0=400000 +y_0=-100000 +ellps=airy +units=km +no_defs" )

### Crop to rectangle
range(st_coordinates(London)[,1])
range(st_coordinates(London)[,2])

all <- subset(all, Easting >= range(st_coordinates(London)[,1])[1] - 10 &
                Easting <= range(st_coordinates(London)[,1])[2]  + 10 &
                Northing >= range(st_coordinates(London)[,2])[1] - 10 &
                Northing <= range(st_coordinates(London)[,2])[2] + 10)

```


Add missing env types and add some additional variables to sitedata 

```{r add}

griddata <- subset(all, is.na(`Site Name`))
sitedata <- subset(all, !is.na(`Site Name`))

missing.sites <- data.frame(`Site Name` = c("London Cromwell Road 2", "London Haringey", "London Haringey Priory Park South", "Southwark A2 Old Kent Road", "Tower Hamlets Roadside", "Thurrock"),
                            `Environment Type` = c("Urban Traffic", "Urban Background", "Urban Background", "Urban Traffic", "Urban Traffic", "Urban Background"))

names(missing.sites) <- c("Site Name", "Environment Type")

all <- left_join(all, missing.sites, by = c("Site Name"))

all$`Environment Type` <- ifelse(!is.na(all$`Environment Type.x`), all$`Environment Type.x`, all$`Environment Type.y`)

names(all)
all <- all %>% select(-c("Environment Type.x", "Environment Type.y"))
unique(all$`Environment Type`)

```

Site types

```{r env}

sitedata <- subset(all, !is.na(`Site Name`))
sites <- unique(sitedata[,c("Site Name","Network","Environment Type")])
count(sites$`Environment Type`)

all$rural <- ifelse(all$`Environment Type` %in% c("Rural Background"), 1, 0)
all$urban <- ifelse(all$`Environment Type` %in% c("Rural Background","Suburban", "Suburban Industrial"), 0, 1) # NOTE: exclusionary
all$suburban <- ifelse(all$`Environment Type` %in% c("Suburban", "Suburban Industrial"), 1, 0)

all$background <- ifelse(all$`Environment Type` %in% c("Kerbside", "Industrial", "Roadside", "Urban Traffic"), 0, 1) # NOTE: exclusionary
all$industrial <- ifelse(all$`Environment Type` %in% c("Industrial", "Urban Industrial", "Suburban Industrial"), 1, 0)
all$traffic <- ifelse(all$`Environment Type` %in% c("Kerbside", "Roadside", "Urban Traffic"), 1, 0)

all$Baseline <- ifelse(all$Network == "AURN", 0, 1)

```


Duplicate site, choose the London Air one

```{r duplicate}

all <- subset(all, is.na(`Site Name`) | !(Network == "AURN" & `Site Name` == "London Teddington Bushy Park"))

sites <- unique(subset(all, !is.na(`Site Name`))[,c("Site Name","Network","Environment Type", "urban","rural","suburban","background","industrial","traffic")])

```



Traffic laws

```{r traffic}

## traffic laws

CC <- c(54:120) #June 2014
all$CC <- as.factor(ifelse(all$Time %in% CC, 1, 0))

LEZ <- c(25:120) #Jan 2012
all$LEZ <- as.factor(ifelse(all$Time %in% LEZ, 1, 0))

TChar <- c(94:101) # Oct 2017 - Apr 2019
all$TChar <- as.factor(ifelse(all$Time %in% TChar, 1, 0))

ULEZ <- c(101:120) # Apr 2019
all$ULEZ <- as.factor(ifelse(all$Time %in% ULEZ, 1, 0))

```


Missing pop

```{r pop}

all$pop <- ifelse(is.na(all$pop), 0, all$pop)

```

Corine 

```{r corine}

all$loc.type <- as.character(factor(all$corine, levels = c("Urban","Rural","Industrial/Transport")))

all$loc.rural <- ifelse(all$loc.type == "Rural", 1, 0)
all$loc.ind <- ifelse(all$loc.type == "Industrial/Transport", 1, 0)

#all %>% count(loc.rural)
#all %>% count(loc.ind)

all$loc.type <- factor(all$loc.type, levels = c("Urban", "Rural", "Industrial/Transport"),
                       labels = c("Urban", "Rural", "Industrial/Transport"))

```

OSM and Corine

```{r osm}

all$URT <- ifelse(all$major50, "Traffic",
                   ifelse(all$minor10, "Traffic", NA))

all$corr.UR <- ifelse(all$LABEL1 %in% c("Agricultural areas", "Forest and semi natural areas", "Water bodies", "Wetlands"),
                         "Rural", "Urban")

all$URT <- ifelse(!is.na(all$URT), all$URT, all$corr.UR)

all$Rural <- ifelse(all$URT == "Rural", 1, 0)
all$Traffic <- ifelse(all$URT == "Traffic", 1, 0)


save(all, file = "DATA/Data/D-All w New Road Dens.RData")

```


Cut time

```{r remove}

load(file = "DATA/Data/D-All w New Road Dens.RData")

unique(all$Time)

all <- subset(all, Time >= 49)
all$Time <- all$Time - 48
all <- subset(all, industrial == 0 | is.na(industrial))

save(all, file = "DATA/Data/FINAL 72 months.RData")

```


```{r road dens}

load(file = "DATA/Data/FINAL 72 months.RData")

all.no2 <- all

all <- subset(all, is.na(`Site Name`) | (PM25 >= 0 & !(`Site Name` %in% c("Kensington and Chelsea - Green Screen BG",
                                                                          "Kensington and Chelsea - Green Screen RS",
                                                                          "Camden - Coopers Lane"))))

save(all, file = "DATA/Data/FINAL PM25 Unscaled.RData")

all <- subset(all.no2, is.na(`Site Name`) | NO2 >= 0)

save(all, file = "DATA/Data/NO2 Unscaled.RData")


```


Scaling

Numeric variables but not scaling:
- Green_Area: 0-100, skewed left
- Land: 400000-1000000, skewed right
- NDVI: -1000-10000, skewed left and out of range
- NDVI.scaled: Scaled, -0.1-1
- NDVI.scaled.valid: 0-1
- smod: not continuous
- Wind deg and dir: Not continuous
- Distance to road: 0-6, skewed left
- Road densities: 0 - 256, bounded


```{r sites}

load(file = "DATA/Data/FINAL 72 months.RData")

all$logPM25PCM <- log(all$PM25PCM)
all$logNO2PCM <- log(all$NO2PCM)
all$logPM25AQUM <- log(all$PM25AQUM)
all$logNO2AQUM <- log(all$NO2AQUM)
all$logPM25MO <- log(all$PM25MO)
all$logNO2MO <- log(all$NO2MO)
all$logPM25MO_2 <- log(all$PM25MO_2)
all$logNO2MO_2 <- log(all$NO2MO_2)
all$logPM25LAEI <- log(all$PM25LAEI)
all$logNO2LAEI <- log(all$NO2LAEI)

all$Green_Area <- all$Green_Area/100

all$primary_dens <- ifelse(is.na(all$primary_dens), 0, all$primary_dens)
all$major_dens <- ifelse(is.na(all$major_dens), 0, all$major_dens)
all$primaryonly_dens <- ifelse(is.na(all$primaryonly_dens), 0, all$primaryonly_dens)
all$minor_dens <- ifelse(is.na(all$minor_dens), 0, all$minor_dens)
all$all_dens <- ifelse(is.na(all$all_dens), 0, all$all_dens)


all$land.scaled <- (all$land - range(all$land)[1])/(range(all$land)[2] - range(all$land)[1])

all$Road_dens.scaled <- (all$Road_dens - range(all$Road_dens)[1])/(range(all$Road_dens)[2] - range(all$Road_dens)[1])

all$major_dens.scaled <- (all$major_dens - range(all$major_dens)[1])/(range(all$major_dens)[2] - range(all$major_dens)[1])

all$primary_dens.scaled <- (all$primary_dens - range(all$primary_dens)[1])/(range(all$primary_dens)[2] - range(all$primary_dens)[1])

all$primaryonly_dens.scaled <- (all$primaryonly_dens - range(all$primaryonly_dens)[1])/(range(all$primaryonly_dens)[2] - range(all$primaryonly_dens)[1])

all$minor_dens.scaled <- (all$minor_dens - range(all$minor_dens)[1])/(range(all$minor_dens)[2] - range(all$minor_dens)[1])

all$all_dens.scaled <- (all$all_dens - range(all$all_dens)[1])/(range(all$all_dens)[2] - range(all$all_dens)[1])



```


```{r scaling}

griddata <- subset(all, is.na(`Site Name`))
sitedata <- subset(all, !is.na(`Site Name`))

names(griddata)[order(names(griddata))]
names(sitedata)[order(names(sitedata))]

num.covars <- c("AOD_adj", "AOD_adj2", "AOD_IDW",  "MaxTemp",
                "BLH", "built_s", "built_v", "Evapo","GHSL.pop","Humid",
               "NO2AQUM", "NO2MO","NO2MO_2","NO2PCM", "NO2LAEI", "PM25MO_2",
                "PM25AQUM", "PM25PCM", "PM25MO","PM25LAEI", "pop","Precip",
                "Press", "Temp","TNO2_IDW", "Wind_speed", "Wind_u","Wind_v")

mean.covars <- apply(griddata[,num.covars], 2, mean, na.rm = TRUE)
sd.covars <- apply(griddata[,num.covars], 2, sd, na.rm = TRUE)

griddata[,num.covars] <- scale(griddata[,num.covars], center = mean.covars, scale = sd.covars)

sitedata[,num.covars] <- scale(sitedata[,num.covars], center = mean.covars, scale = sd.covars)

```

And for road distance

```{r road}

max.maj <- range(griddata$dist.to.maj)[2]
max.min <- range(griddata$dist.to.min)[2]
max.road <- range(griddata$dist.to.road)[2]

# scaling to 0-1

max.km <- max(max.maj, max.min, max.road)
griddata[,c("dist.to.maj", "dist.to.min", "dist.to.road")] <- griddata[,c("dist.to.maj", "dist.to.min", "dist.to.road")]/max.km

# scaling sitedata on the same scale

range(sitedata$dist.to.maj)
range(sitedata$dist.to.min)
range(sitedata$dist.to.road)

sitedata[,c("dist.to.maj", "dist.to.min", "dist.to.road")] <- sitedata[,c("dist.to.maj", "dist.to.min", "dist.to.road")]/max.km


all <- rbind(griddata, sitedata)




```

Save 

```{r save}

save(all, file = "DATA/Data/FINAL ALL.RData")

```
