---
title: "1-AURN Redownload"
author: "Abi Riley"
date: "2024-05-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "Z:/home")
```

## Processing AURN monitoring site data

Load in packages

```{r packages, echo=FALSE}

library(readxl)
library(ggplot2)
library(reshape2)
library(scico)
library(dplyr)
library(tidyr)
library(xts)
library(rquery)

```


## Load in data

Loading in monitoring site monthly means and monitoring site information

```{r data}

NO2 <- read.csv("Jobs/ST_INLA/FINAL/FINAL/Full Data Processing/New Datasets/Data/AURN_NO2.csv", header = FALSE)

PM25 <- read.csv("Jobs/ST_INLA/FINAL/FINAL/Full Data Processing/New Datasets/Data/AURN_PM25.csv", header = FALSE)

AURNSites <- read_excel("Objective 1/Data/AURN Monitoring Sites/All AURN MS.xlsx")

```

# Remove unwanted info

```{r data2}
no2.names <- NO2[5,]
NO2 <- NO2[-c(1:5),]
names(NO2) <- no2.names
NO2 <- subset(NO2, Units == "ugm-3")

pm25.names <- PM25[5,]
PM25 <- PM25[-c(1:5),]
names(PM25) <- pm25.names
PM25 <- subset(PM25, Units == "ugm-3")


```


Set variable name and make as long dataframes

```{r make_long}
NO2 <- NO2 %>%
  pivot_longer(!`Site Name`:Year, names_to = "Month", values_to = "Monthly Mean")
PM25 <- PM25 %>%
  pivot_longer(!`Site Name`:Year, names_to = "Month", values_to = "Monthly Mean")

NO2 <- subset(NO2, Month != "Units")
PM25 <- subset(PM25, Month != "Units")

names(NO2)[10] <- "NO2"
names(PM25)[10] <- "PM25"


```

Merge PM$_{2.5}$ and NO$_2$ dataframes

```{r merge}

All <- natural_join(NO2, PM25, by =  c("Year","Month", "Site Name"), jointype = "FULL")

All$Latitude <- as.numeric(All$Latitude)
All$Longitude <- as.numeric(All$Longitude)

All$NO2 <- as.numeric(All$NO2)
All$PM25 <- as.numeric(All$PM25)

```

Merge with monitoring site data

```{r add_info}

All <- natural_join(All, AURNSites, by = "Site Name", jointype = "LEFT")

```

Make time variables

```{r dates}

All$month <- as.numeric(factor(All$Month, levels = c("Jan","Feb","Mar","Apr","May","Jun","Jul","Aug","Sep","Oct","Nov","Dec")))

All$Year <- as.numeric(All$Year)

All$Time <- (All$Year - 2010)*12 + All$month

```

Make environment type varaibles for industrial and traffic

```{r env_type}

All$industrial <- ifelse(grepl("Industrial", All$`Environment Type`, fixed = TRUE), 1, 0)
All$traffic <- ifelse(grepl("Traffic", All$`Environment Type`, fixed = TRUE), 1, 0)

```

Order by time and site and save dataframe

```{r save}
All <- All[order(All$Time),]
All <- All[order(All$`Site Name`),]

save(All, file = "Z:/home/DATA/Data/1-AURN.RData")

```