---
title: "15-Corine"
author: "Abi Riley"
date: "2024-10-29"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "Z:/home")

```

## Packages 

```{r packages, echo= FALSE}

library(terra)
library(readxl)
library(ggplot2)
library(dplyr)
library(tidyr)
library(sf)

```

Read in data

```{r data}

corine <- read_sf("Z:/home/Objective 1/Data/Corine/data/clc2018_uk.shp")

legend <- read_excel("Z:/home/Objective 1/Data/Corine/supporting-documents/clc_legend.xls")

```


Project

``` {r crop}

TTW <- read_sf("Z:/home/Jobs/ST_INLA/Shapefiles/Travel_to_Work_Areas_(Dec_2011)_FEB_in_United_Kingdom.shp")
London2 <- subset(TTW, OBJECTID == 92 | OBJECTID == 122)
London2 <- st_transform(London2, crs = "EPSG:27700" )

corine <- st_transform(corine, crs = "EPSG:27700")

```

Rasterise

``` {r rasterize}
bbox <- rast(ncol=88, nrow=90, 
             xmin=500000, xmax=588000,
             ymin=135000, ymax=225000,
             crs = "EPSG:27700")


st_bbox(corine)

names(corine)

corine.r <- rast(corine, resolution = 1000)
corine.z <- rasterize(corine, corine.r, field = "CODE_18")
corine.london <- project(corine.z, bbox)
corine.df <- as.data.frame(corine.london, xy = TRUE)


plot(corine.london)

```



Merge with legend and make categories

```{r merge}

names(legend)[2] <- "CODE_18"
corine <- left_join(corine.df, legend, by = "CODE_18")

View(unique(corine[,c("LABEL1","LABEL2","LABEL3")]))

corine$corrine <- ifelse(corine$LABEL1 %in% c("Agricultural areas", "Forest and semi natural areas"), "Rural",
                         ifelse(corine$LABEL1 %in% c("Water bodies","Wetlands"), "Water/Wetlands",
                                ifelse(corine$LABEL2 == "Urban fabric", "Urban",
                                       ifelse(corine$LABEL2 %in% c("Industrial, commercial and transport units","Mine, dump and construction sites"), "Industrial/Transport", 
                                              ifelse(corine$LABEL3 == "Sport and leisure facilities", "Suburban", 
                                                     ifelse(corine$LABEL3 == "Green urban areas", "Urban", NA))))))


corine$corrine <- ifelse(corine$LABEL1 %in% c("Agricultural areas", "Forest and semi natural areas"), "Rural",
                         ifelse(corine$LABEL1 %in% c("Water bodies","Wetlands"), "Rural",
                                ifelse(corine$LABEL2 == "Urban fabric", "Urban",
                                       ifelse(corine$LABEL2 %in% c("Industrial, commercial and transport units","Mine, dump and construction sites"), "Industrial/Transport", 
                                              ifelse(corine$LABEL3 == "Sport and leisure facilities", "Urban", 
                                                     ifelse(corine$LABEL3 == "Green urban areas", "Urban", NA))))))

ggplot() + 
  geom_raster(corine, mapping = aes(x = x, y = y, fill = corrine))

```


LABEL1: Agricultural areas -> Rural

LABEL1: Artificial surfaces -
          - Urban fabric (Discontinuous and continuous urban fabric) -> Urban
          
          - Artificial, non-agricultural vegetated areas (Sport and leisure facilities) -> Suburban
          - Artificial, non-agricultural vegetated areas (Green urban areas) -> Urban
          
          - Industrial, commercial and transport units (Industrial or commercial units) -> Industrial/Transport
          - Industrial, commercial and transport units (Road and rail networks and associated land) -> Industrial/Transport
          - Industrial, commercial and transport units (Airports) -> Industrial/Transport
          - Industrial, commercial and transport units (Port areas) -> Industrial/Transport
          - Mine, dump and construction sites (Mineral extraction sites) -> Industrial/Transport
          - Mine, dump and construction sites (Construction sites) -> Industrial/Transport
          - Mine, dump and construction sites (Dump sites) -> Industrial/Transport
          
LABEL1: Forest and semi natural areas -> Rural
LABEL1: Water bodies -> Water/Wetlands
LABEL1: Wetlands -> Water/Wetlands

         

```{r save}


save(corine, file = "DATA/Data/15-Corine.RData")


```
