---
title: "D-Roads"
author: "Abi Riley"
date: "2024-10-31"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "Z:/home")

```

https://jcoliver.github.io/learn-r/017-open-street-map.html

## Packages 

```{r packages, echo= FALSE}

library(osmdata)
library(ggplot2)
library(sf)
library(terra)

```

Data

```{r data}

load(file = "DATA/Data/C-All and Sitedata b.RData")

sitedata <- subset(all, !is.na(`Site Name`))
all.grid <- subset(all, is.na(`Site Name`))


```

Transform to Lat/Long

```{r lat/long}

all.sf <- st_as_sf(all.grid, coords = c("Easting","Northing"), crs = "+proj=tmerc +lat_0=49 +lon_0=-2 +k=0.999601272 +x_0=400000 +y_0=-100000 +ellps=airy +units=m +no_defs")
all.sf <- st_transform(all.sf, crs = "EPSG:4326")
st_bbox(all.sf)

```

London bounding box

```{r bbox}

London <- matrix(data = c(-0.5302177, 0.3660678, 51.2656611, 51.7113983),
                 nrow = 2,
                 byrow = TRUE)
colnames(London) <- c("min", "max")
rownames(London) <- c("x", "y")

```

Load in roads from osm

```{r osm-roads}

London_motorway <- getbb(place_name = "London") %>%
  opq(timeout = 50) %>%
  add_osm_feature(key = "highway", 
                  value = c("motorway")) %>%
  osmdata_sf()

London_motorway_sf <- st_as_sf(London_motorway$osm_lines) %>% 
  st_transform(crs = "+proj=tmerc +lat_0=49 +lon_0=-2 +k=0.999601272 +x_0=400000 +y_0=-100000 +ellps=airy +units=km +no_defs")
st_write(London_motorway_sf$geometry, "DATA/Data/D-OSM Motorways 3.shp", append = FALSE)

London_primary <- getbb(place_name = "London") %>%
  opq(timeout = 50) %>%
  add_osm_feature(key = "highway", 
                  value = c("motorway", "primary")) %>%
  osmdata_sf()

London_primary_sf <- st_as_sf(London_primary$osm_lines) %>% 
  st_transform(crs = "+proj=tmerc +lat_0=49 +lon_0=-2 +k=0.999601272 +x_0=400000 +y_0=-100000 +ellps=airy +units=km +no_defs")
st_write(London_primary_sf$geometry, "DATA/Data/D-OSM Primary Roads 2.shp", append = FALSE)

London_primary_only <- getbb(place_name = "London") %>%
  opq(timeout = 50) %>%
  add_osm_feature(key = "highway", 
                  value = c("primary")) %>%
  osmdata_sf()

London_primary_only_sf <- st_as_sf(London_primary_only$osm_lines) %>% 
  st_transform(crs = "+proj=tmerc +lat_0=49 +lon_0=-2 +k=0.999601272 +x_0=400000 +y_0=-100000 +ellps=airy +units=km +no_defs")
st_write(London_primary_only_sf$geometry, "DATA/Data/D-OSM Primary Roads Only.shp", append = FALSE)


load(file = "DATA/Data/D-OSM Major Roads.RData")
major.lines <- London_major$osm_lines %>% 
  st_transform(crs = "+proj=tmerc +lat_0=49 +lon_0=-2 +k=0.999601272 +x_0=400000 +y_0=-100000 +ellps=airy +units=km +no_defs")
st_write(major.lines$geometry, "DATA/Data/D-OSM Major Roads 2.shp", append = FALSE)

load(file = "DATA/Data/D-OSM Minor Roads.RData")
minor.lines <- London_minor$osm_lines %>% 
  st_transform(crs = "+proj=tmerc +lat_0=49 +lon_0=-2 +k=0.999601272 +x_0=400000 +y_0=-100000 +ellps=airy +units=km +no_defs")
st_write(minor.lines$geometry, "DATA/Data/D-OSM Minor Roads.shp", append = FALSE)


```


Method:

1. Take centroid (Easting + 0.5km and Northing + 0.5km)
2. First, major road within 50m radii
3. Second, minor road within 10m radii
4. Else, corrine urban/rural classification
5. Check again site types and normal corrine



Grid centroids

```{r grid}

x <- rep(seq(502.5, 563.5, 1), 49)
y.coords <- rep(seq(154.5, 202.5, 1), 62)
y <- y.coords[order(y.coords)]
grid <- as.data.frame(cbind(x, y))
head(grid)

grid$grid.id <- 1:nrow(grid)

grid.r <- terra::rast(grid, type = "xyz", 
                      crs = "+proj=tmerc +lat_0=49 +lon_0=-2 +k=0.999601272 +x_0=400000 +y_0=-100000 +ellps=airy +units=km +no_defs")



```

Site data to sf

```{r sites}

sites <- st_drop_geometry(unique(sitedata[,c("Site Name", "Easting", "Northing")]))
sites$siteid <- 1:nrow(sites)
sites[,c("Easting","Northing")] <- sites[,c("Easting","Northing")]/1000

sitedata[,c("Easting","Northing")] <- sitedata[,c("Easting","Northing")]/100

sites.sf <- st_as_sf(sites, coords = c("Easting","Northing"), crs = "+proj=tmerc +lat_0=49 +lon_0=-2 +k=0.999601272 +x_0=400000 +y_0=-100000 +ellps=airy +units=km +no_defs")


```



```{r 50m}

# major
major.lines <- London_major$osm_lines
major.sf <- st_as_sf(major.lines)
major.sf <- st_transform(major.sf, crs = "+proj=tmerc +lat_0=49 +lon_0=-2 +k=0.999601272 +x_0=400000 +y_0=-100000 +ellps=airy +units=km +no_defs")

major.v <- as.lines(vect(major.sf))
grid.major <- terra::extract(grid.r, major.v)
plot(grid.major)

grid.major.df <- left_join(grid, grid.major, by = "grid.id") 
grid.major.count <- count(grid.major.df, "grid.id")

grid.wMajor <- left_join(grid, grid.major.count, by = "grid.id")
grid.wMajor$major <- ifelse(is.na(grid.wMajor$freq), 0,
                            ifelse(grid.wMajor$freq >= 1, 1, 0))

#primary

primary.lines <- London_primary$osm_lines
primary.sf <- st_as_sf(primary.lines)
primary.sf <- st_transform(primary.sf, crs = "+proj=tmerc +lat_0=49 +lon_0=-2 +k=0.999601272 +x_0=400000 +y_0=-100000 +ellps=airy +units=km +no_defs")

primary.v <- as.lines(vect(primary.sf))
grid.primary <- terra::extract(grid.r, primary.v)
plot(grid.primary)

grid.primary.df <- left_join(grid, grid.primary, by = "grid.id") 
grid.primary.count <- count(grid.primary.df, "grid.id")

grid.wprimary <- left_join(grid, grid.primary.count, by = "grid.id")
grid.wprimary$primary <- ifelse(is.na(grid.wprimary$freq), 0,
                            ifelse(grid.wprimary$freq >= 1, 1, 0))

#motorway
motorway.lines <- London_motorway$osm_lines
motorway.sf <- st_as_sf(motorway.lines)
motorway.sf <- st_transform(motorway.sf, crs = "+proj=tmerc +lat_0=49 +lon_0=-2 +k=0.999601272 +x_0=400000 +y_0=-100000 +ellps=airy +units=km +no_defs")

motorway.v <- as.lines(vect(motorway.sf))
grid.motorway <- terra::extract(grid.r, motorway.v)
plot(grid.motorway)

grid.motorway.df <- left_join(grid.motorway, grid, by = "grid.id")
grid.motorway.count <- count(grid.motorway.df, "grid.id")

grid.wMotorway <- left_join(grid, grid.motorway.count, by = "grid.id")


grid.wMotorway$motorway <- ifelse(is.na(grid.wMotorway$freq), 0,
                            ifelse(grid.wMotorway$freq >= 1, 1, 0)) 

grid.wMotorway <- grid.wMotorway %>% 
  select(-c("freq"))

save(grid.wMotorway, file = "DATA/Data/Motorway_Indicator.RData")

motorway.r <- rast(grid.wMotorway, type = "xyz")
plot(motorway.r)

# with 50m buffer
motorway.lines2 <- st_transform(motorway.lines, crs = "+proj=tmerc +lat_0=49 +lon_0=-2 +k=0.999601272 +x_0=400000 +y_0=-100000 +ellps=airy +units=km +no_defs") %>% 
  select(c("osm_id", "ref", "geometry"))

grid.df <- crds(grid.r, df = TRUE)
grid.sf <- st_as_sf(crds(grid.r, df = TRUE), coords = c("x", "y"), crs = "+proj=tmerc +lat_0=49 +lon_0=-2 +k=0.999601272 +x_0=400000 +y_0=-100000 +ellps=airy +units=km +no_defs")

grid.buffer <- sf::st_buffer(grid.sf,
              dist = 1.05,
              endCapStyle = "SQUARE")

grid.buffer$ID <- 1:nrow(grid.buffer)

grid.motorway.df <- as.data.frame(st_intersects(grid.buffer, motorway.lines2))

names(grid.motorway.df)[1] <- "grid.id"

grid.motorway.df <- left_join(grid.motorway.df, grid, by = "grid.id")
grid.motorway.count <- count(grid.motorway.df, "grid.id")

grid.wMotorway <- left_join(grid.wMotorway, grid.motorway.count, by = "grid.id")

grid.wMotorway$motorway.buffer <- ifelse(is.na(grid.wMotorway$freq), 0,
                            ifelse(grid.wMotorway$freq >= 1, 1, 0)) 


save(grid.wMotorway, file = "DATA/Data/Motorway_Indicator_wBuffer.RData")

grid.wMotorway <- grid.wMotorway[,c("x", "y", "motorway", "motorway.buffer")]
names(grid.wMotorway)[1:2] <- c("Easting", "Northing")

save(grid.wMotorway, file = "DATA/Data/Motorway_Indicator_wBuffer_Named.RData")


```


```{r}
### which motorways

motorway.sf$ID <- 1:nrow(motorway.sf)
motorways <- unique(motorway.sf[,c("ID", "ref")])
grid.motorway.names <- left_join(grid.motorway.df, motorways, by = "ID")
grid.motorway.names$motorway <- 1

grid.motorway.names.wide <- spread(grid.motorway.names, ref, motorway)
grid.motorway.names.wide$motorways <- print()

library(stringr)
grid.motorway.names.wide$motorways <- apply(grid.motorway.names.wide[6:14], 1, function(x) toString(names(x)[x >= 1]))



```


Traffic indicator

```{r traffic}

grid$URT <- ifelse(grid$major50, "Traffic",
                   ifelse(grid$minor10, "Traffic", NA))
sites$URT <- ifelse(sites$major50, "Traffic",
                    ifelse(sites$minor10, "Traffic", NA))

```

Join

```{r join}

load(file = "DATA/Data/C-All and Sitedata b.RData")

sitedata <- subset(all, !is.na(`Site Name`))
sitedata[,c("Easting","Northing")] <- sitedata[,c("Easting","Northing")]/1000

sites <- sites[,c("Site Name","major50","minor10","URT")]
sitedata <- left_join(sitedata, sites, by = "Site Name")

all.grid <- subset(all, is.na(`Site Name`))
all.grid[,c("Easting","Northing")] <- all.grid[,c("Easting","Northing")]/1000

names(grid)[c(1,2)] <- c("Easting", "Northing")
all <- left_join(all.grid, grid, by = c("Easting", "Northing"))



```

Add corine to label Urban-Rural

*LABEL1: Agricultural areas -> Rural
*LABEL1: Artificial surfaces -
*   - Urban fabric (Discontinuous and continuous urban fabric) -> Urban
*   - Artificial, non-agricultural vegetated areas (Sport and leisure facilities) -> Urban?
*   - Artificial, non-agricultural vegetated areas (Green urban areas) -> Urban
*   - Industrial, commercial and transport units -> Urban
*   - Mine, dump and construction sites -> Urban?
* LABEL1: Forest and semi natural areas -> Rural
* LABEL1: Water bodies -> Rural
* LABEL1: Wetlands -> Rural

```{r corine}

all$corr.UR <- ifelse(all$LABEL1 %in% c("Agricultural areas", "Forest and semi natural areas", "Water bodies", "Wetlands"),
                         "Rural", "Urban")
sitedata$corr.UR <- ifelse(sitedata$LABEL1 %in% c("Agricultural areas", "Forest and semi natural areas", "Water bodies", "Wetlands"),
                      "Rural", "Urban")

all$URT <- ifelse(!is.na(all$URT), all$URT, all$corr.UR)
sitedata$URT <- ifelse(!is.na(sitedata$URT), sitedata$URT, sitedata$corr.UR)


```

Save

```{r save2}

save(all, file = "DATA/Data/D-All b.RData")
save(sitedata, file = "DATA/Data/D-Sitedata b.RData")

```


Add new roads

OSM sum of road lengths

```{r lengths}

load(file = "DATA/Data/D-All b.RData")
load(file = "DATA/Data/D-Sitedata b.RData")

major_density <- read.csv("DATA/Data/QGIS/Major_density2.csv") 
primary_density <- read.csv("DATA/Data/QGIS/MotorwayPrimary_density2.csv")
motorway_density <- read.csv("DATA/Data/QGIS/Motorway_density2.csv")
primaryonly_density <- read.csv("DATA/Data/QGIS/PrimaryOnly_density.csv")


road_density <- cbind(major_density, primary_density$COUNT, motorway_density$COUNT, primaryonly_density$COUNT)
names(road_density)[c(10:13)] <- c("major_dens", "primary_dens", "motorway_dens","primaryonly_dens")

road_density$Easting <- (road_density$left + 500)/1000
road_density$Northing <- (road_density$bottom + 500)/1000

road_density <- road_density[,c("Easting", "Northing", "major_dens", "primary_dens", "motorway_dens", "primaryonly_dens")]

save(road_density, file = "DATA/Data/Road_density_OSM.RData")


```
