---
title: "19-Roads"
author: "Abi Riley"
date: "2024-12-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "Z:/home")

```


## Packages 

```{r packages, echo= FALSE}

library(spatstat)
library(sp)
library(gstat)
library(sf)
library(dplyr)
library(MatrixModels)
library(gtools)
library(rquery)
library(terra)
library(fields)

```


Get site coordinates

```{r coords}

load(file = "DATA/Data/3-AURN and LA Data.RData")

sitedata <- subset(sitedata, Easting >= 502000 & Easting <= 564000 & Northing >= 153000 & Northing <= 203000)

sites <- sitedata %>% select(c("Site Name","Easting", "Northing")) %>% unique()

sites[,c("Easting", "Northing")] <- sites[,c("Easting", "Northing")]/1000

sites$siteid <- 1:nrow(sites)
 
sites.sf <- st_as_sf(sites, coords = c("Easting","Northing"), crs = "+proj=tmerc +lat_0=49 +lon_0=-2 +k=0.999601272 +x_0=400000 +y_0=-100000 +ellps=airy +units=km +no_defs")

```

Make grid 

```{r grid}

x <- rep(seq(502.5, 563.5, 1), 49)
y.coords <- rep(seq(154.5, 202.5, 1), 62)
y <- y.coords[order(y.coords)]
grid <- as.data.frame(cbind(x, y))
head(grid)

grid$grid.id <- 1:nrow(grid)

grid.sf <- st_as_sf(grid, coords = c("x","y"), crs = "+proj=tmerc +lat_0=49 +lon_0=-2 +k=0.999601272 +x_0=400000 +y_0=-100000 +ellps=airy +units=km +no_defs")

```
London bounding box

```{r bbox}

London <- matrix(data = c(-0.5302177, 0.3660678, 51.2656611, 51.7113983),
                 nrow = 2,
                 byrow = TRUE)
colnames(London) <- c("min", "max")
rownames(London) <- c("x", "y")

```

OSM getting major roads

"highway": "motorway", "primary", "secondary"

```{r major}

available_tags(feature = "highway")

London_major <- getbb(place_name = "London") %>%
  opq(timeout = 50) %>%
  add_osm_feature(key = "highway", 
                  value = c("motorway", "primary", "secondary")) %>%
  osmdata_sf()

London_major

street_plot <- ggplot() +
  geom_sf(data = London_major$osm_lines,
          inherit.aes = FALSE,
          color = "black",
          size = 0.2)
street_plot

```

OSM getting minor roads

"highway": "tertiary", "residential"

```{r minor}

street_plot <- ggplot() +
  geom_sf(data = London_major$osm_lines,
          inherit.aes = FALSE,
          color = "black",
          size = 0.2)
street_plot

London_minor <- getbb(place_name = "London") %>%
  opq() %>%
  add_osm_feature(key = "highway", value = c("tertiary", "residential")) %>%
  osmdata_sf()

street_plot <- street_plot +
  geom_sf(data = London_minor$osm_lines,
          inherit.aes = FALSE,
          color = "#666666",  # medium gray
          size = 0.1) # half the width of the major roads
street_plot

```

Save plot and two datasets


```{r save}

ggsave("DATA/Data/19-OSM Roads.pdf")

save(London_major, file = "DATA/Data/19-OSM Major Roads.RData")

save(London_minor, file = "DATA/Data/19-OSM Minor Roads.RData")

```

Load

```{r load}

load(file = "DATA/Data/19-OSM Major Roads.RData")

load(file = "DATA/Data/19-OSM Minor Roads.RData")

```


Method:

1. Take centroid (Easting + 0.5km and Northing + 0.5km)
2. First, major road within 50m radii
3. Second, minor road within 10m radii
4. Else, corrine urban/rural classification
5. Check again site types and normal corrine

50m buffer to major road for grid and sites


```{r 50m}

grid.50 <- st_buffer(grid.sf, dist = 0.05)
sites.50 <- st_buffer(sites.sf, dist = 0.05)

major.lines <- London_major$osm_lines
major.sf <- st_as_sf(major.lines)
major.sf <- st_transform(major.sf, crs = "+proj=tmerc +lat_0=49 +lon_0=-2 +k=0.999601272 +x_0=400000 +y_0=-100000 +ellps=airy +units=km +no_defs")

grid.major <- unlist(st_intersects(major.sf, grid.50))
grid$major50 <- ifelse(grid$grid.id %in% grid.major, 1, 0)
grid %>% count(major50)

sites.major <- unlist(st_intersects(major.sf, sites.50))
sites$major50 <- ifelse(sites$siteid %in% sites.major, 1, 0)
sites %>% count(major50)

```


10m buffer to minor road for grid and sites

```{r 10m}

grid.10 <- st_buffer(grid.sf, dist = 0.01)
sites.10 <- st_buffer(sites.sf, dist = 0.01)

minor.lines <- London_minor$osm_lines
minor.sf <- st_as_sf(minor.lines, crs = "EPSG:4326")
minor.sf <- st_transform(minor.sf, crs = "+proj=tmerc +lat_0=49 +lon_0=-2 +k=0.999601272 +x_0=400000 +y_0=-100000 +ellps=airy +units=km +no_defs")

grid.minor <- unlist(st_intersects(minor.sf, grid.10))
grid$minor10 <- ifelse(grid$grid.id %in% grid.minor, 1, 0)
grid %>% count(minor10)

sites.minor <- unlist(st_intersects(minor.sf, sites.10))
sites$minor10 <- ifelse(sites$siteid %in% sites.minor, 1, 0)
sites %>% count(minor10)

```

Save

```{r save}

save(grid, file = "DATA/Data/19-Grid Major Minor Roads.RData")
save(sites, file = "DATA/Data/19-Sites Major Minor Roads.RData")

```

Load

```{r save}

load(file = "DATA/Data/19-Grid Major Minor Roads.RData")
load(file = "DATA/Data/19-Sites Major Minor Roads.RData")

```

Can add traffic indicator when I do corrine


# Distance to Road

## Distance of the nearest road to the centroid/site

min distance of point to line

```{r distance}

major.lines <- London_major$osm_lines
major.sf <- st_as_sf(major.lines)
major.sf <- st_transform(major.sf, crs = "+proj=tmerc +lat_0=49 +lon_0=-2 +k=0.999601272 +x_0=400000 +y_0=-100000 +ellps=airy +units=km +no_defs")

minor.lines <- London_minor$osm_lines
minor.sf <- st_as_sf(minor.lines, crs = "EPSG:4326")
minor.sf <- st_transform(minor.sf, crs = "+proj=tmerc +lat_0=49 +lon_0=-2 +k=0.999601272 +x_0=400000 +y_0=-100000 +ellps=airy +units=km +no_defs")

```

Sitedata 

```{r sitedata}

sites.sf <-  st_as_sf(sites, coords = c("Easting","Northing"), crs = "+proj=tmerc +lat_0=49 +lon_0=-2 +k=0.999601272 +x_0=400000 +y_0=-100000 +ellps=airy +units=km +no_defs")

sites.sf[,c("dist.to.maj", "dist.to.min", "dist.to.road")] <- NA

site <- list()
for(i in 1:nrow(sites.sf)){
  site[[i]] <- sites.sf[i,]
  site[[i]]$dist.to.maj <- min(st_distance(site[[i]], major.sf))
  site[[i]]$dist.to.min <- min(st_distance(site[[i]], minor.sf))
  
  site[[i]]$dist.to.road <- min(site[[i]]$dist.to.maj, site[[i]]$dist.to.min)
}

sites <- do.call(rbind, site)

save(sites, file = "DATA/Data/19-Sitedata Dist to Road.RData")


```

Grid

```{r gridcells}

grid.sf <- st_as_sf(grid, coords = c("x","y"), crs = "+proj=tmerc +lat_0=49 +lon_0=-2 +k=0.999601272 +x_0=400000 +y_0=-100000 +ellps=airy +units=km +no_defs")

grid.sf[,c("dist.to.maj", "dist.to.min", "dist.to.road")] <- NA

grid <- list()
for(i in 1:nrow(grid.sf)){
  grid[[i]] <- grid.sf[i,]
  grid[[i]]$dist.to.maj <- min(st_distance(grid[[i]], major.sf))
  grid[[i]]$dist.to.min <- min(st_distance(grid[[i]], minor.sf))
  
  grid[[i]]$dist.to.road <- min(grid[[i]]$dist.to.maj, grid[[i]]$dist.to.min)
}

grids <- do.call(rbind, grid)

save(grids, file = "DATA/Data/19-Grid Dist to Road.RData")

```


```{r check}

load(file = "DATA/Data/19-Sitedata Dist to Road.RData")
load(file = "DATA/Data/19-Grid Dist to Road.RData")

summary(sites[,c("dist.to.maj", "dist.to.min", "dist.to.road")])
summary(grids[,c("dist.to.maj", "dist.to.min", "dist.to.road")])


```