---
title: "Correlation Analysis"
author: "Abi Riley"
date: "2025-05-13"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "Z:/home")
```

## Processing AURN monitoring site data

Load in packages

```{r packages, echo=FALSE}

library(sp)
library(INLA)
library(gstat)
library(devtools)
library(sf)
library(inlabru)
library(ggplot2)
library(tidyr)
library(dplyr)
library(corrplot)
library(pheatmap)

```


## Load in data

```{r data}

load(file = "DATA/Data/FINAL ALL.RData")

all <- subset(all, is.na(`Site Name`) | PM25 >= 0)
sitedata <- subset(all,  PM25 >= 0)

names(sitedata)[order(names(sitedata))]


```

# Correlation between variables

## All Variables

```{r all}

data <- sitedata[,c("PM25", "logPM25", "Time",   
                    "all_dens.scaled","AOD_adj2",                  
                    "background", "Baseline",                    
                    "built_s", "built_v",                
                    "dist.to.maj", "dist.to.min", "dist.to.road",          
                    "Easting", "Evapo",                  
                    "GHSL.pop", "Green_Area", "Humid",                  
                    "land.scaled",                    
                    "loc.ind", "loc.rural",               
                    "major_dens.scaled", "major50", "MaxTemp", "minor_dens",             
                    "minor_dens.scaled", "minor10", "NDVI_New",               
                    "PM25AQUM", "PM25LAEI", "PM25MO_2", "PM25PCM",                
                    "PM25Sat", "pop", "Precip","Press",  
                    "primary_dens.scaled","primaryonly_dens.scaled",                   
                    "Road_dens.scaled", "Rural", "suburban", "Temp",                 
                    "traffic", "Traffic",                
                    "urban", "Wind_speed","Wind_u", "Wind_v")]

corr <- cor(data, use = "pairwise.complete.obs")
cor(data)[1,]

```


## Corrplot using the package 'corrplot' to create plot 

```{r plot}

corrplot(corr, method = "color", order = "FPC", type = "lower", 
         tl.col = "black", tl.srt = 45)


cor(data, use = "pairwise.complete.obs")[1,]


```

Or with numbers instead of colours:

```{r plot2}

corrplot(corr, method = 'number')


```

## A clustered heatmap with the package 'pheatmap'

This clusters variables with similar correlations and that explain similar processes.

```{r pheatmap}

pheatmap::pheatmap(corr, scale = "row")

```

## Hierarchical Clustering 

We can also create hierarchical clustering based on distances.

    Distance = 1 - Correlation

Then the function 'hclust' to plot a correlation matrix with dendograms of the different layers of correlation.

We can plot just the dendogram and we can cut the dendogram at different k groups


```{r plot}

dist <- as.dist(1 - corr)
tree <- hclust(dist, method = "complete")
plot(tree)
dend <- as.dendrogram(tree)
plot(dend)
cutree(tree, k = 5)


```


# Correlation by month

We can explore how correlation between the variable and PM$_{2.5}$ change by month.

We select just the continuous variables

```{r continuous}

continuous.covariates <- c("all_dens.scaled","AOD_adj2",                  
                "Baseline",                    
                "built_s", "built_v",                
                "dist.to.maj", "dist.to.min", "dist.to.road",          
                "Easting", "Evapo",                  
                "GHSL.pop", "Green_Area", "Humid",                  
                "land.scaled", "loc.ind", "loc.rural",               
                "major_dens.scaled","MaxTemp",             
                "minor_dens.scaled", "NDVI_New",               
                "PM25MO_2", "PM25PCM",                
                "pop", "Precip","Press",  
                "primary_dens.scaled",                   
                "Road_dens.scaled", "Temp",                 
                "Wind_speed","Wind_u", "Wind_v")

```

Using a loop over each variable, calculate and print the correlation between PM$_{2.5}$ and the variable each calendar month. Plot for each variable.

```{r corr-month}

par(mfrow=c(4,8))
for(j in 1:length(continuous.covariates)){
  cor <- c()
  for(i in 1:12){
    var <- continuous.covariates[j]
    sitedata.month <- subset(sitedata, month == i)
    cor[i] <- cor(sitedata.month[,c(paste(var), "logPM25")], use = "pairwise.complete.obs")[1,2]
  }
  plot(1:12, cor, type = "l", main = var)
  print(var)
  #print(range(cor))
}


```

We can also look at the 95\% range of the correlation values by calendar month.

```{r corr-month-95}

par(mfrow=c(4,8))
for(j in 1:length(continuous.covariates)){
  cor <- c()
  for(i in 1:12){
    var <- continuous.covariates[j]
    sitedata.month <- subset(sitedata, month == i)
    cor[i] <- cor(sitedata.month[,c(paste(var), "logPM25")], use = "pairwise.complete.obs")[1,2]
  }
  plot(1:12, cor, type = "l", main = var)
  print(var)
  #print(range(cor))
  print(quantile(cor, 0.025))
  print(quantile(cor, 0.975))
}

```
