---
title: "Spike-and-Slab Prior"
author: "Abi Riley"
date: "2025-05-13"
output: html_document
---


```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "Z:/home")
set.seed(23)


```

Load in packages

```{r packages, echo=FALSE}

library(sp)
library(INLA)
library(gstat)
library(devtools)
library(sf)
library(inlabru)
library(ggplot2)
library(tidyr)
library(dplyr)
library(BoomSpikeSlab)

```


## Load in data

```{r data}

load(file = "DATA/Data/FINAL ALL.RData")

all <- subset(all, is.na(`Site Name`) | PM25 >= 0)
sitedata <- subset(all,  PM25 >= 0)

names(sitedata)[order(names(sitedata))]

```

## Select Model Covariates

```{r covariates}


covariates <- c("background", "urban", "Baseline",
                "major50", "minor10", "Traffic",
                "built_s", "built_v", "GHSL.pop",
                "Green_Area", "pop", 
                "dist.to.maj", "dist.to.min", "dist.to.road",
                "Road_dens", "loc.ind", "loc.rural",
                "all_dens.scaled", "major_dens.scaled", 
                "minor_dens.scaled","primary_dens.scaled",
                "PM25MO_2", "PM25PCM",
                "AOD_adj2", "NDVI_New",
                "Humid", "Temp", "MaxTemp", "Precip", "Press",
                "Evapo", "Wind_speed", "Wind_u", "Wind_v")

sitedata$Intercept <- 1

```


# Spike-and-Slab Prior Regression for Variable Selection

## Define model formula with beta spike-and-slab priors on covariates with:
  1. Mean Beta = 0
  2. SD Beta = 1
  3. Theta initial value = -3

```{r formula}

f.iid <- logPM25 ~ -1  +  Intercept  +  
  f(beta, model = "ssvs", mean.beta = 0, prec.beta = 1, hyper = list(theta = list(initial = -3)))

```

## Create matrix of variables

```{r X}

X <- model.matrix(~ all_dens.scaled + AOD_adj2 + background + Baseline + built_s + built_v +
                    dist.to.maj + dist.to.min + dist.to.road + Easting + Evapo + GHSL.pop +
                    Green_Area + Humid + land.scaled + loc.ind + loc.rural + major_dens.scaled +
                    major50 + MaxTemp + minor_dens + minor_dens.scaled + minor10 + NDVI_New +
                    PM25AQUM + PM25LAEI + PM25MO_2 + PM25PCM + pop + Precip + Press +
                    primary_dens.scaled + Road_dens.scaled + Rural + suburban + Temp + TNO2_IDW +
                    traffic + Traffic + urban + Wind_speed + Wind_u + Wind_v,
                  data = sitedata)


```

## Define prior with
  1. Expected number of non-zero coefficients = 2
  2. Prior degrees of freedom = 1 
        - degrees of freedom for estimating the residual variance.
  3. Prior information weight = 0.01 
        - Number of observations worth of weight that should be given to the prior estimate of beta
  4. Diagonal shrinkage = 0
        - Weight given to the diagonal of the average of diagonals of the precision matrix
        
        
```{r prior}

prior <- SpikeSlabPrior(X, sitedata$logPM25,
                        expected.model.size = 2,  
                        prior.df = 1,
                        prior.information.weight = 0.01,
                        diagonal.shrinkage = 0)


```

## Run model and get summary

```{r model}

model <- lm.spike(logPM25 ~ -1 + Intercept + background + urban + built_v + pop + dist.to.min + Road_dens + all_dens +
                    minor_dens + PM25MO_2 + PM25PCM + AOD_adj2 + NDVI_New + Temp + Humid + Press + Evapo +
                    Wind_speed + Wind_u + Wind_v,
                  data = sitedata, niter = 1000, prior = prior)
summary(model) 


```