---
title: "Stepwise Forward Selection"
author: "Abi Riley"
date: "2025-05-13"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "Z:/home")
set.seed(23)

```


## Load in data

```{r data}
load(file = "DATA/Data/FINAL ALL.RData")

all <- subset(all, is.na(`Site Name`) | PM25 >= 0)
sitedata <- subset(all,  PM25 >= 0)

names(sitedata)[order(names(sitedata))]
```

## Add site ID and select considered variables

```{r vars}

sitedata$siteid <- as.numeric(as.factor(paste(sitedata$`Site Name`, sitedata$Network)))
names(sitedata)
covars <- sitedata[,c("background", "urban", "Baseline", 
                          "major50", "minor10", "Traffic",
                          "built_s", "built_v", "GHSL.pop",
                          "Green_Area", "pop", 
                          "dist.to.maj", "dist.to.min", "dist.to.road",
                          "Road_dens", "loc.ind", "loc.rural",
                          "all_dens.scaled", "major_dens.scaled", 
                          "minor_dens.scaled","primary_dens.scaled",
                          "PM25MO_2", "PM25PCM",
                          "AOD_adj2", "NDVI_New",
                          "Humid", "Temp", "MaxTemp", "Precip", "Press",
                          "Evapo", "Wind_speed", "Wind_u", "Wind_v")]


outcome <- "logPM25"

```

## Define 

  1. Predictive model choice criterion (PMCC) 
  2. Model fit correlation
  3. 'write.excel' file function

```{r define}

r2.fit <- function(z, zhat){
  z <- as.matrix(z)
  zhat <- as.matrix(zhat)
  res <- c(z - zhat)
  r2.fitted <- cor(z,zhat)^2
  return(r2.fit)
}


PMCC.fit <- function(z, zhat){
  z <- as.matrix(z)
  zhat <- as.matrix(zhat)
  res <- c(z - zhat)
  gof <- sum(res^2)
  penalty <-  sum(abs(res))
  PMCC.fit <- gof + penalty
  return(PMCC.fit)
}
write.excel <- function(x,row.names=FALSE,col.names=TRUE,...) {
  write.table(x,"clipboard",sep="\t",row.names=row.names,col.names=col.names,...)
}

```

# Stepwise Forward Model Selection

## Select candidate covariates

```{r candidates}

candidate_covariates <- c("background", "urban", "Baseline", 
                          "major50", "minor10", "Traffic",
                          "built_s", "built_v", "GHSL.pop",
                          "Green_Area", "pop", 
                          "dist.to.maj", "dist.to.min", "dist.to.road",
                          "Road_dens", "loc.ind", "loc.rural",
                          "all_dens.scaled", "major_dens.scaled", 
                          "minor_dens.scaled","primary_dens.scaled",
                          "PM25MO_2", "PM25PCM",
                          "AOD_adj2", "NDVI_New",
                          "Humid", "Temp", "MaxTemp", "Precip", "Press",
                          "Evapo", "Wind_speed", "Wind_u", "Wind_v")

sitedata$Intercept <- 1


```

## Loop for running model runs for each additional variable and caluclate the PMCC 

Round 1:
  logPM25 = Intercept

```{r Round1}

results_list <- list()
pmcc <- c()
for(i in 1:(length(candidate_covariates))){
  formula_test <-  reformulate(c("Intercept", 
                                 paste(candidate_covariates[i],collapse="+")), response ="logPM25")
  
  
  # Fit the model
  mod.res <- INLA::inla(formula = formula_test,
                        family = "gaussian",
                        data = sitedata,
                        control.predictor=list(link=1,compute=T),
                        control.compute=list(dic=T, cpo=T, waic = T),
                        control.inla =list(int.strategy = "eb")
  )
  
  # Extract the fitted values
  M_fit <- mod.res$summary.fitted.values[,"mean"]
  
  # R2: Pseudo_r2 function defined above
  r2 <- cor(sitedata$logPM25, mod.res$summary.fitted$mean)^2
  
  # PMCC: PMCC function defined above
  PMCC <- PMCC.fit(sitedata$logPM25, mod.res$summary.fitted$mean)
  
  # Store the results
  results_list[[i]] = list("cov" = candidate_covariates[i],
                           "PMCC" = PMCC)
}
results <- do.call(rbind, results_list)
results
write.excel(results)


```


Round 2:
  logPM25 = Intercept + PM25MO

```{r Round2}

results_list <- list()
pmcc <- c()
for(i in 1:(length(candidate_covariates))){
  formula_test <-  reformulate(c("Intercept", "PM25MO_2",
                                 paste(candidate_covariates[i],collapse="+")), response ="logPM25")
  
  
  # Fit the model
  mod.res <- INLA::inla(formula = formula_test,
                        family = "gaussian",
                        data = sitedata,
                        control.predictor=list(link=1,compute=T),
                        control.compute=list(dic=T, cpo=T, waic = T),
                        control.inla =list(int.strategy = "eb")
  )
  
  # Extract the fitted values
  M_fit <- mod.res$summary.fitted.values[,"mean"]
  
  # R2: Pseudo_r2 function defined above
  r2 <- cor(sitedata$logPM25, mod.res$summary.fitted$mean)^2
  
  # PMCC: PMCC function defined above
  PMCC <- PMCC.fit(sitedata$logPM25, mod.res$summary.fitted$mean)
  
  # Store the results
  results_list[[i]] = list("cov" = candidate_covariates[i],
                           "PMCC" = PMCC)
}
results <- do.call(rbind, results_list)
results
write.excel(results)


```


Round 3:
  logPM25 = Intercept + PM25MO + background

```{r Round3}

results_list <- list()
pmcc <- c()
for(i in 1:(length(candidate_covariates))){
  formula_test <-  reformulate(c("Intercept", "PM25MO_2", "background", "minor_dens.scaled",
                                 paste(candidate_covariates[i],collapse="+")), response ="logPM25")
  
  
  # Fit the model
  mod.res <- INLA::inla(formula = formula_test,
                        family = "gaussian",
                        data = sitedata,
                        control.predictor=list(link=1,compute=T),
                        control.compute=list(dic=T, cpo=T, waic = T),
                        control.inla =list(int.strategy = "eb")
  )
  
  # Extract the fitted values
  M_fit <- mod.res$summary.fitted.values[,"mean"]
  
  # R2: Pseudo_r2 function defined above
  r2 <- cor(sitedata$logPM25, mod.res$summary.fitted$mean)^2
  
  # PMCC: PMCC function defined above
  PMCC <- PMCC.fit(sitedata$logPM25, mod.res$summary.fitted$mean)
  
  # Store the results
  results_list[[i]] = list("cov" = candidate_covariates[i],
                           "PMCC" = PMCC)
}
results <- do.call(rbind, results_list)
results
write.excel(results)


```


Round 4:
  logPM25 = Intercept + PM25MO + background + minor_dens

```{r Round4}

results_list <- list()
pmcc <- c()
for(i in 1:(length(candidate_covariates))){
  formula_test <-  reformulate(c("Intercept", "PM25MO_2", "background", "minor_dens.scaled",
                                 paste(candidate_covariates[i],collapse="+")), response ="logPM25")
  
  
  # Fit the model
  mod.res <- INLA::inla(formula = formula_test,
                        family = "gaussian",
                        data = sitedata,
                        control.predictor=list(link=1,compute=T),
                        control.compute=list(dic=T, cpo=T, waic = T),
                        control.inla =list(int.strategy = "eb")
  )
  
  # Extract the fitted values
  M_fit <- mod.res$summary.fitted.values[,"mean"]
  
  # R2: Pseudo_r2 function defined above
  r2 <- cor(sitedata$logPM25, mod.res$summary.fitted$mean)^2
  
  # PMCC: PMCC function defined above
  PMCC <- PMCC.fit(sitedata$logPM25, mod.res$summary.fitted$mean)
  
  # Store the results
  results_list[[i]] = list("cov" = candidate_covariates[i],
                           "PMCC" = PMCC)
}
results <- do.call(rbind, results_list)
results
write.excel(results)


```


Round 5:
  logPM25 = Intercept + PM25MO + background + minor_dens + Traffic

```{r Round4}

results_list <- list()
pmcc <- c()
for(i in 1:(length(candidate_covariates))){
  formula_test <-  reformulate(c("Intercept", "PM25MO_2", "background", "minor_dens.scaled", 
                                 "Traffic",
                                 paste(candidate_covariates[i],collapse="+")), response ="logPM25")
  
  
  # Fit the model
  mod.res <- INLA::inla(formula = formula_test,
                        family = "gaussian",
                        data = sitedata,
                        control.predictor=list(link=1,compute=T),
                        control.compute=list(dic=T, cpo=T, waic = T),
                        control.inla =list(int.strategy = "eb")
  )
  
  # Extract the fitted values
  M_fit <- mod.res$summary.fitted.values[,"mean"]
  
  # R2: Pseudo_r2 function defined above
  r2 <- cor(sitedata$logPM25, mod.res$summary.fitted$mean)^2
  
  # PMCC: PMCC function defined above
  PMCC <- PMCC.fit(sitedata$logPM25, mod.res$summary.fitted$mean)
  
  # Store the results
  results_list[[i]] = list("cov" = candidate_covariates[i],
                           "PMCC" = PMCC)
}
results <- do.call(rbind, results_list)
results
write.excel(results)


```


END

Chosen variables are:
    1. Met Office AQR PM2.5 Model
    2. Background site indicator
    3. Density of minor roads from OSM
    4. Traffic environment indicator from OSM 
