---
title: "VIF and Autocorrelation"
author: "Abi Riley"
date: "2025-05-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "Z:/home")
set.seed(23)


```

Load in packages

```{r packages, echo=FALSE}

library(sp)
library(INLA)
library(gstat)
library(devtools)
library(sf)
library(inlabru)
library(ggplot2)
library(tidyr)
library(dplyr)
library(BoomSpikeSlab)

```


## Load in data

```{r data}

load(file = "DATA/Data/FINAL ALL.RData")

all <- subset(all, is.na(`Site Name`) | PM25 >= 0)
sitedata <- subset(all,  PM25 >= 0)

names(sitedata)[order(names(sitedata))]

```

## Select Model Covariates

```{r covariates}


covariates.names <- c("background", "urban", "Baseline",
                "major50", "minor10", "Traffic",
                "built_s", "built_v", "GHSL.pop",
                "Green_Area", "pop", 
                #"dist.to.maj", "dist.to.min", "dist.to.road",
                "Road_dens", "loc.ind", "loc.rural",
                "all_dens.scaled", "major_dens.scaled", 
                #"minor_dens.scaled","primary_dens.scaled",
                "PM25MO_2", "PM25PCM",
                "AOD_adj2", "NDVI_New",
                "Humid", "Temp", "MaxTemp", "Precip", "Press",
                "Evapo", "Wind_speed", "Wind_u", "Wind_v")

sitedata$Intercept <- 1

```

# Perform Covariate Selection

Using the package 'spatialRF', we:
    1. Set thresholds for autocorrelation and VIF
    2. Calculate the autocorrelation 'auto_cor'
    3. Cacluate the VIF
    4. Use the autocorrelation and VIF thresholds to include/exclude variables
    
    $VIF_i = \frac{1}{1-R_i^2}$ 
    
    for each variable i
    
```{r covar_select}

new.covariates <- spatialRF::auto_cor(
  x = sitedata[, covariates.names],
  cor.threshold = 0.9,
) %>% 
  spatialRF::auto_vif(
    vif.threshold = 1.2,
)

new.covariates$selected.variables


```

Or in a loop

```{r loop}

cor.set <- c(0.3, 0.4, 0.5, 0.6, 0.7, 0.8)
vif.set <- c(1.2, 1.5, 2, 2.5)

new.covariates <- data.frame(cor.set = NA, vif.set = NA)
new.covariates[,c(3:20)] <- NA

for(i in 1:length(cor.set)){
  for(j in 1:length(vif.set)){
    
    cor.threshold <-  cor.set[i]
    vif.threshold <- vif.set[j]
  
    variables <- spatialRF::auto_cor(
      x = sitedata[, covariates.names],
      cor.threshold = cor.threshold
      ) %>%
    
    spatialRF::auto_vif(
      vif.threshold = vif.threshold
      )

    new.variables <- c(cor.threshold, vif.threshold, c(variables$selected.variables))
    
    if(length(variables$selected.variables) < 18){
      new.variables[(length(variables$selected.variables) + 2):20] <- NA
    } else {
      new.variables <- new.variables
    }
    
    new.covariates <- rbind(new.covariates, new.variables)
    
  }
}

new.covariates <- new.covariates[-1,]


````


## Count occurance of each variable

```{r count}

counts <- table(unlist(new.covariates[-c(1,2)]))

counts[order(counts)]

```